---
title: "Text Mining of Tweets of Trump"
author: "Fei Shi"
date: "December 18, 2017"
output:
  beamer_presentation: default
  ioslides_presentation:
    fig_height: 2.9
    fig_width: 5.4
  slidy_presentation: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = FALSE, message = FALSE, comment = "")
```


##Introduction

After June 6, 2016,  Hillary Clinton became the presumptive Democratic nominee, and after that Trump began campaigning against Hillary Clinton, we want to know what changes of Donald Trump before and after knowing campaigning against Hillary Clinton, is he some feared or disappointed? 


##Material and Methods

The data are tweets from Donald John Trump range  from 2015-12-14 to 2016-08-08.  There are 1512 with 16 variables in the original data, we only use variables 'id', 'created', 'text' in our analysis:

- 'id' is the uniqute identifier for each tweet
- 'created' is the time line 
- 'text' is the content of the tweet

##Tools


- reference from the text book https://www.tidytextmining.com/ as well as github https://github.com/dgrtwo/dgrtwo.github.com/blob/master/_R/.

- r packages

```{r,echo=TRUE}
library(dplyr)
library(tidyr)
library(lubridate)
library(scales)
library(ggplot2)
library(stringr)
library(tidytext)
library(broom)
library(wordcloud)

```


##Time series



```{r}
#load(url("http://varianceexplained.org/files/trump_tweets_df.rda"))
load("trump_tweets_df.rda")
tweets <- trump_tweets_df %>%
  select(id,  text, created) %>%
  mutate(Status = ifelse(created < "2016-6-6", "Before", "After"))
tweets %>%
  count(Status, hour = hour(with_tz(created, "EST"))) %>%
  mutate(percent = n / sum(n)) %>%
  ggplot(aes(hour, percent, color = Status)) +
  geom_line() +   labs(x = "Hour of day (EST)",
       y = "% of tweets",
       color = "")
```

Trump becomes much more busy while having much less time to post tweets after knowing campaigning against Hillary Clinton.


## Words comparisions


```{r}
reg <- "([^A-Za-z\\d#@']|'(?![A-Za-z\\d#@]))"
tweet_words <- tweets %>%
  filter(!str_detect(text, '^"')) %>%
  mutate(text = str_replace_all(text, "https://t.co/[A-Za-z\\d]+|&amp;", "")) %>%
  unnest_tokens(word, text, token = "regex", pattern = reg) %>%
  filter(!word %in% stop_words$word,
         str_detect(word, "[a-z]"))

tweet_words %>%
  count(word, sort = TRUE) %>%
  head(20) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n)) +
  geom_bar(stat = "identity") +
  ylab("Occurrences") +
  coord_flip()
```

We remove tweets which are likely not be posted by Trump himself, Obviously, he care more about hillary cliton than himself.

## Words comparisions


```{r}
beofre_after_ratios <- tweet_words %>%
  count(word, Status) %>%
  filter(sum(n) >= 5) %>%
  spread(Status, n, fill = 0) %>%
  ungroup() %>% 
  mutate(logratio = log2((Before +1) / (After+1))) %>%
  arrange(desc(logratio))

beofre_after_ratios %>%
  group_by(logratio > 0) %>%
  top_n(15, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  ylab("Before / after log ratio") 
```

The words most different from two status  he cares about 'vote trump' and 'debate' as well as 'woman' after compaign agaist Hillary clition. 


##wordclouds

```{r, fig.width=7.8}
frequency <- tweet_words %>% 
  group_by(Status) %>% 
  count(word, sort = TRUE) %>% 
  left_join(tweet_words %>% 
              group_by(Status) %>% 
              summarise(total = n())) %>%
  mutate(freq = n/total)


frequency <- frequency %>% 
  select(Status, word, freq) %>% 
  spread(Status, freq) %>%
  arrange(Before, After)


temp =  tweet_words %>% group_by(Status) %>% 
  anti_join(stop_words) %>%
  count(word)

par(mfrow = c(1,2))
temp2 = temp %>% filter(Status == "Before") %>% 
  with(wordcloud(word, n, max.words = 100))
temp3 = temp %>% filter(Status == "After")%>% 
  with(wordcloud(word, n, max.words = 100))
```
 
 Trump is selfish before  but more care hillary, clinton after 2016-06-06.

##scatter plot
 
```{r}
ggplot(frequency, aes(Before, After)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

Trump cares about job and people also voters as the three words are in the line, also, Trump always like joke, so joke is shown in the line too, on the top of the tail of the line, we can find hillary clinton is in the side of after, it means Trump indeed talks about hillary clinton more.

##Sentiment analysis



```{r}
nrc <- sentiments %>%
  filter(lexicon == "nrc") %>%
  dplyr::select(word, sentiment)

```


```{r}
sources <- tweet_words %>%
  group_by(Status) %>%
  mutate(total_words = n()) %>%
  ungroup() %>%
  distinct(id, Status, total_words)



by_source_sentiment <- tweet_words %>%
  inner_join(nrc, by = "word") %>%
  count(sentiment, id) %>%
  ungroup() %>%
  complete(sentiment, id, fill = list(n = 0)) %>%
  inner_join(sources) %>%
  group_by(Status, sentiment, total_words) %>%
  summarize(words = sum(n)) %>%
  ungroup()

```




```{r,fig.height=3.2}

sentiment_differences <- by_source_sentiment %>%
  group_by(sentiment) %>%
  do(tidy(poisson.test(.$words, .$total_words)))


sentiment_differences %>%
  ungroup() %>%
  mutate(sentiment = reorder(sentiment, estimate)) %>%
  mutate_each(funs(. - 1), estimate, conf.low, conf.high) %>%
  ggplot(aes(estimate, sentiment)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
  scale_x_continuous(labels = percent_format()) +
  labs(x = "% increase in after relative to before",
       y = "Sentiment")
```

Use Exact Poisson tests to test whether the ratio between two rate parameters are equal or not as well as compute confidence intervals.

##Sentiment analysis


-  fear, anger, negative, disgust, sadness are the main sentiments after campaigning against Hillary Clinton

on one hand, Trump might fear failture, on the the hand, he defame Hillary Clinton. However, he sometimes feel anger and sadness about rumors from Hillary Clinton. 



